---
title: "Additional Exercises"
---

```{r, echo = FALSE, purl = FALSE, message = FALSE}
source("setup.R")
```

# Session 1 - 
Extra exercises inspired by "Getting started" & "Introduction to R programming".  

> *Challenge*.  
> Recall that we suggested that you create a "standard" directory structure. Write a code snippet or script to create your preferred structure as a utility to be run when you create a new Rstudio project.   
> Consider how you could test if one or more of the directories already exists in your current working directory.   
> __Hint:__ You may find it useful to checkout Rs help on __'files2'__ and __'getwd'__.   
>
> ```{r, answer=TRUE, purl=FALSE}
> curr_dir <- getwd()
> print(curr_dir)
> if (!dir.exists("data"))  dir.create("data")
>
> # Another (and more flexible ) way could use the 'lapply' function (see R help) like this:    
> tst_apply <- c("d1","d2","d3") # a vector of sub-dirs to create
> lapply(tst_apply, dir.create)
> ```

*Challenge*.  
Using two of the functions we introduced in this session, find out how many missing variables are missing in the cereals dataset. \
NB You will need to execute "install.packages("plspm")" first to install package and its dataset.   
```{r}
# Install dataset
library(plspm)
data("cereals")
# read in surveys as dataframe (assumes portal data in data dir)
surveys <- read.csv("data/portal_data_joined.csv")

```

> ```{r, answer=TRUE, purl=FALSE}
> # we can add up the number of TRUEs using the sum() function
> sum(is.na(cereals))
> # if this is zero then there are no missing values.  
> ```

Work through the surveys data-set and find which columns have missing values and how many.    
For an extra challenge use the help (or google) to use a __for__ loop to count the number of NAs in each column and the __cat()__ to print out a table.   
```{r}
sum(is.na(surveys))
coln <- colnames(surveys)
length(coln)
coln
# could print sum(is.na(surveys, 1)) 2 etc to print
# tota NAs for each column or....
out_lst <- list()
for( i in 1: length(coln))
  cat(i, coln[i], sum(is.na(surveys[,i])),"\n")
```

# Session 2 -
Extra exercises inspired by "Starting with data".   
The data set contains 77 cereals with 15 variables:

manufacturer (categorical).  
type (categorical).  
number of calories (numeric).  
grams of protein (numeric).  
grams of fat (numeric).  
milligrams of sodium (numeric).  
grams of dietary fiber (numeric).  
grams of complex carbohydrates (numeric).  
grams of sugars (numeric).   
grams of potassium (numeric).  
percentage of FDA recommended vitamins (numeric).   
display shelf (categorical).  
weight in ounces (numeric).  
number of cups in one serving (numeric).  
rating of the cereal (numeric).  

```{r}
summary(cereals)
```

*Challenge*.  
* List data for the cereals manufactured by firm 'K'.   
```{r}
cereals[cereals$mfr == "K",]
```

* Find the first ten cereals that have greater than 100 calories per serving.   
```{r}
head(cereals[cereals$calories > 100,],10)
```

* Of the cereals that have greater than 100 calories per serving, what is the mean number of calories/serving?
```{r}
mean(cereals$calories[cereals$calories > 100])
```

* List those cereals made by 'K' with less than 100 calories per serving.   
```{r}
cereals[cereals$calories < 100 & cereals$mfr == "K",]
```

# Session 3 -
Extra exercises inspired by session 1 of  "Data manipulation and visualisation with tidyverse".  

Load 'tidyverse' and convert cereals to a tibble (either by saving to CSV and then reading back into a tibble or by learning to use the __as_tibble__ function from R help).   
```{r}
library(tidyverse)
cereals_t <- as_tibble(cereals)
```

*Challenge*.  

It is hypothesised that consumers prefer cereals that contain more sugar. Plot the relationship between sugar content and rating.
Does this support the hypothesis?   

```{r}
ggplot(data = cereals_t,mapping = aes(x = sugars, y = rating)) +
  geom_point() +
  geom_smooth()
```

It is possible that market dominance is skewing our sample.   
Produce a bar-plot showing the number of items made by each manufacturer in our data-set.   
```{r}
ggplot(data=cereals_t,mapping = aes(x=mfr)) +
 geom_bar()
```

In section 2 you listed those cereals made by 'K' with less than 100 calories per serving using the 'Base R' method. Please produce a Tidyverse version (without creating any intermediate variables or nesting functions). List the manufacturer, calories and cereal name.   
```{r}
cereals_t %>%
  filter(calories < 100 & mfr == "K") %>%
  select(mfr,calories, type)
```


# Session 4 -
Extra exercises inspired by session 2 of "Data manipulation and visualisation with tidyverse".  

*Challenge*.  

# Capstone project (Putting it all together) -
