---
title: "Additional Exercises"
---

```{r, echo = FALSE, purl = FALSE, message = FALSE}
source("setup.R")
```

### Extended activities
```{r, echo=FALSE, purl=TRUE, warning=FALSE}
```

# Session 1

Extra exercises inspired by *"Getting started" & "Introduction to R programming"*.

> *Challenge*.\
> Recall that we suggested that you create a "standard" directory structure. Write a code snippet or script to create your preferred structure as a utility to be run when you create a new Rstudio project.\
> Consider how you could test if one or more of the directories already exists in your current working directory.\
> **Hint:** You may find it useful to checkout Rs help on **'files2'** and **'getwd'**.
>
> <details>
>
> <summary>Answer</summary>
>
> ```{r, answer=TRUE, purl=FALSE}
> curr_dir <- getwd()
> print(curr_dir)
> if (!dir.exists("data"))  dir.create("data")
> >
> # Another (and more flexible ) way could use the 'lapply' function (see R help) like this:    
> tst_apply <- c("d1","d2","d3") # a vector of sub-dirs to create
> lapply(tst_apply, dir.create)
> ```
>
> </details>

> #### *Challenge*.
>
> Using two of the functions we introduced in this session, find out how many missing variables are missing in the cereals dataset.\
> NB You will need to execute "install.packages("plspm")" first to install package and its dataset.

> ```{r}
> # Install dataset
> library(plspm)
> data("cereals")
> # read in surveys as dataframe (assumes portal data in data dir)
> surveys <- read.csv("data/portal_data_joined.csv")
> ```
>
> <details>
>
> <summary>Answer</summary>
>
> ```{r, answer=TRUE, purl=FALSE}
> # we can add up the number of TRUEs using the sum() function
> sum(is.na(cereals))
> # if this is zero then there are no missing values.  
> ```
>
> </details>

> Work through the **surveys** data-set and find which columns have missing values and how many.\
> For an extra challenge use the help (or google) to use a **for** loop to count the number of NAs in each column and the **cat()** to print out a table.
>
> <details>
>
> <summary>Answer</summary>
>
> ```{r}
> sum(is.na(surveys))
> coln <- colnames(surveys)
> length(coln)
> coln
> # could print sum(is.na(surveys, 1)) 2 etc to print
> # tota NAs for each column or....
> out_lst <- list()
> for( i in 1: length(coln))
>   cat(i, coln[i], sum(is.na(surveys[,i])),"\n")
> ```
>
> </details>

# Session 2

Extra exercises inspired by *"Starting with data"*.\
The data set contains 77 cereals with 15 variables:

manufacturer (categorical)\
\* A = American Home Food Products\
\* G = General Mills\
\* K = Kelloggs\
\* N = Nabisco\
\* P = Post\
\* Q = Quaker Oats\
\* R = Ralston Purina  

type (categorical).\
\* C = cold cereal\
\* H = hot cereal\
*Nutritional information per serving*.\
number of calories (numeric).\
grams of protein (numeric).\
grams of fat (numeric).\
milligrams of sodium (numeric).\
grams of dietary fiber (numeric).\
grams of complex carbohydrates (numeric).\
grams of sugars (numeric).\
grams of potassium (numeric).\
percentage of FDA recommended vitamins (numeric).\
display shelf (categorical) 1=bottom shelf to 3=top.\
weight in ounces (numeric).\
number of cups in one serving (numeric).\
rating of the cereal (numeric) - consumer survey?.

```{r}
summary(cereals)
```

*Challenges*.

-   List data for the cereals manufactured by firm 'K'.

<details>

<summary>Answer</summary>

> ```{r}
> cereals[cereals$mfr == "K",]
> ```

</details>

.

-   Find the first ten cereals that have greater than 100 calories per serving.

<details>

<summary>Answer</summary>

```{r}
head(cereals[cereals$calories > 100,],10)
```

</details>

-   Of the cereals that have greater than 100 calories per serving, what is the mean number of calories/serving?

<details>

<summary>Answer</summary>

```{r}
mean(cereals$calories[cereals$calories > 100])
```

</details>

-   List those cereals made by 'K' with less than 100 calories per serving.

<details>

<summary>Answer</summary>

```{r}
cereals[cereals$calories < 100 & cereals$mfr == "K",]
```

</details>

.

# Session 3

Extra exercises inspired by session 1 of *"Data manipulation and visualisation with tidyverse"*.

Load 'tidyverse' and convert cereals to a tibble (either by saving to CSV and then reading back into a tibble or by learning to use the **as_tibble** function from R help).\

<details>

<summary>Answer</summary>

```{r}
library(tidyverse)
cereals_t <- as_tibble(cereals)
```

</details>

*Challenge*.\
It is hypothesised that consumers prefer cereals that contain more sugar. Plot the relationship between sugar content and rating. Does this support the hypothesis?

<details>

<summary>Answer</summary>

```{r, warning=FALSE}
# challenge phrasing suggests using facets
ggplot(data = cereals_t,mapping = aes(x = sugars, y = rating)) +
  geom_point() +
  geom_smooth()
```

</details>

It is possible that market dominance is skewing our sample.\
Produce a bar-plot showing the number of items made by each manufacturer in our data-set.\

<details>

<summary>Answer</summary>

```{r}
ggplot(data=cereals_t,mapping = aes(x=mfr)) +
 geom_bar()
```

</details>

In section 2 you listed those cereals made by 'K' with less than 100 calories per serving using the 'Base R' method. Please produce a Tidyverse version (without creating any intermediate variables or nesting functions). List the manufacturer, calories and cereal name.\

<details>

<summary>Answer</summary>

```{r}
cereals_t %>%
  filter(calories < 100 & mfr == "K") %>%
  select(mfr,calories, type)
```

</details>

# Session 4

Extra exercises inspired by session 2 of *"Data manipulation and visualisation with tidyverse"*.

*Challenges*.\
Using Tidyverse functions, create a summary table of the average calories by manufacturer.\
It should look like this:  

| Manufacturer | mean_calories |
| :---: | :---:|
| A	| 100.00000	|
| G	| 111.36364	|	
| K	| 108.69565		|
| N	| 86.66667	|	
| P	| 108.88889	|	
| Q	| 95.00000		|
| R	| 115.00000	|.  

<details>

<summary>Answer</summary>

```{r}
# Look up rename() in help - it's a useful dplyr verb
cereals_t %>%
  group_by(mfr) %>%
  rename(Manufacturer = mfr) %>%
  summarise(mean_calories=mean(calories))
```

</details>

Produce a scatterplot of calories vs. sugar grouped by manufacturer.   

<details>
<summary>Answer</summary>
```{r,warning=FALSE, message=FALSE}
ggplot(data = cereals_t, mapping = aes(x = sugars, y = calories)) +
    geom_point() +
    geom_smooth() +
    facet_wrap(facets = vars(mfr))
```

</details>

<hr>

# Capstone project (Putting it all together).

This will be performed on a dataset from a paper by [Reaven & Miller - downloadable from here](https://web.math.utk.edu/~fernando/Students/GregClark/pdf/Miller-Reaven-Diabetes.pdf). This research analyses the stages of the disease Type 2 Diabetes.\
You will need to install the 'heplots' package and load it using library() and this will bring in the data-set we will be working with.\

<details>

<summary>Answer</summary>

```{r}
# list dataset
library(heplots)
Diabetes
```

</details>

## Gain familiarity with the Diabetes dataset.

We have 5 columns with numeric values (relwt, glufast, glutest, instest, sspg) and one categorical column (group).\
Details about these values are to be found [here](https://search.r-project.org/CRAN/refmans/heplots/html/Diabetes.html).

## Exploring the dataset.

```{r}
print(summary(Diabetes))
```

From this we find that Group comprises of 3 values: (Normal, Chemical_Diabetic, Overt_Diabetic) corresponding to stages of developing Type 2 diabetes.

## Histogram plots of the data.

Use the Base R **hist()** command (Consult R help for usage) to plot the distributions of the numeric variables - are they normally distributed (Bell curve)?.

<details>

<summary>Answer</summary>

```{r, warning=FALSE}
# Plot histograms of all columns distributions
# Does data fit normality criteria for statistical tests?
# e.g. if we wanted to do a t-test. We'll just use base R histogam
# here - As an extra exercise: How would you do this in tidyverse
# as well?
hist(Diabetes$relwt)  
hist(Diabetes$glufast)
hist(Diabetes$glutest)  
hist(Diabetes$instest)
hist(Diabetes$sspg)
```

</details>

## Explore correlations.

Install the "PerformanceAnalytics" package to help us create correlation plots for our data-set.

NB.\
- This plot also provides another way of generating the histograms of the data columns.\
- The distribution of each variable is shown on the diagonal.\
- On the bottom of the diagonal : the bivariate scatter plots with a fitted line are displayed.\
- On the top of the diagonal : the value of the correlation plus the significance level as stars.\
- Each significance level is associated to a symbol :\
p-values(0, 0.001, 0.01, 0.05, 0.1, 1) \<=\> symbols(“***”, “**”, “*”, “.”, " “).

<details>

<summary>Answer</summary>

```{r, warning=FALSE, message=FALSE}
# install.packages("PerformanceAnalytics")
library("PerformanceAnalytics")
my_data <- Diabetes[, c(1,2,3,4,5)]
chart.Correlation(my_data, histogram=TRUE, pch=19)
```

</details>
